---
output:
  pdf_document: default
  html_document: default
---
# Chemicals that Drive Aroma Ratings



```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
```

We want to remove the effect of the subjects from our response, this will give us better results for
chemicals that drive aroma scores. We will residualize the response by fitting a model for 
Aroma ~ Subject variables, then find the residuals, we will then use the residuals as our response 
and implement 3 things to screen and test screening ElasticNet, SIS and PLS. We will also use ML models
like XGBoost to find the important variables but I'd be very concerned about using them because of the
size of our dataset.

```{r include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(glmnet)
```

```{r include=FALSE}
subject_df <- read.csv("Subject_Level_Dairy_Cheese_latent_factors.csv")
df_chem <- read.csv("Cheese_Chemical_Data.csv")
rating_df <- read.csv("Subject_wise_Cheese_Ratings.csv")
```

Variables to consider for subject level data: "Frequency_of_PB_cheese", "Diet", "Age", 
"Cheese_MR1", "Cheese_MR2".

Data Prep:

Chemical Data:
```{r}
chem_cols <- colnames(df_chem)[4:54]

chem_screen_df <- df_chem %>%
  select(all_of(c("PrimaryID", chem_cols)))



#Log-transforming the chemical data
chem_screen_df[, 2:52] <- log(chem_screen_df[, 2:52])

X_chem_scaled <- chem_screen_df

X_chem_scaled[, chem_cols] = chem_screen_df %>%
  select(all_of(chem_cols)) %>%
  scale() 
head(X_chem_scaled)
```


Ratings and Subject Data:
```{r}
subject_vars <- c("Frequency_of_PB_cheese", "Diet", 
                "Age", "Cheese_MR1", "Cheese_MR2")

analysis_df <- rating_df %>%
  left_join(subject_df, by = "Unique_Panelist_ID")

needed_cols <- c("Unique_Panelist_ID", "PrimaryID", "Aroma_A", subject_vars)

analysis_df <- analysis_df %>%
  select(all_of(needed_cols))

analysis_df$Diet <- factor(analysis_df$Diet)

analysis_df$Age <- factor(analysis_df$Age, ordered = T)
analysis_df$Unique_Panelist_ID <- factor(analysis_df$Unique_Panelist_ID)
```


Alternate: One big joint model


Mixed effects model with a random intercept for subject:

```{r}
m_subject <- lmer(data = analysis_df, 
                formula = Aroma_A ~ (1|Unique_Panelist_ID)+ 
                Frequency_of_PB_cheese + Age + Diet + Cheese_MR1 + Cheese_MR2,
                REML = TRUE)

#Residualizing the response
analysis_df$resid_subject <- resid(m_subject)

summary(m_subject)
```


Cheese level data for response:

```{r}
cheese_target_df <- analysis_df %>%
  group_by(PrimaryID) %>%
  summarise(
    mean_rating = mean(Aroma_A),
    mean_resid_subject = mean(resid_subject),
    n_ratings = n(),
    .groups = "drop"
  )
names(cheese_target_df)[1] <- "PrimaryID"

write.csv(cheese_target_df, 
"cheese_level_residualized_target.csv", row.names = FALSE)
```


Merging residualized response with cheese level data (chemicals):

```{r}
screen_df <- cheese_target_df %>%
            left_join(X_chem_scaled, by = "PrimaryID")

screen_df[!complete.cases(screen_df),]

rbind(head(screen_df), tail(screen_df))
```


Now we will move to the first screening method using Elastic Net (this will shrink collinear 
and unimportant chemicals coefficient to 0).
(Elastic Net has L1 and L2 Regularization terms)
Elastic net's penalty adds the following to the usual loss function of linear regression:
$$ g(\beta) = \lambda [\alpha ||\beta||_1 + \frac{1-\alpha}{2} ||\beta||_2^2] $$


```{r}
set.seed(123)

X <- screen_df %>%
  select(all_of(chem_cols)) %>%
  as.matrix()

y <- screen_df[["mean_resid_subject"]]

cv_enet <- cv.glmnet(
  x = X,
  y = y,
  alpha = 0.5,          # elastic net
  family = "gaussian",
  standardize = TRUE,
  nfolds = min(10, nrow(X))
)
plot(cv_enet)
```

lambda is the regularization term, picking between -log(lambda) between 2.5 and 3 seems reasonable,
bigger lambda's seem to have diminishing returns and provide a more complex model.
Let's consider -log(lambda) = 2.8, i.e. lambda = exp(-2.8), we will compare this to the optimal 
lambda from cross validation and check which gives us better results, I'd like a simpler model.

```{r}
lambda_custom <- exp(-2.8)
lambda_custom
```

Comparing Elastic Net models:
```{r}
coef_min <- coef(cv_enet, s = "lambda.min")
coef_1se <- coef(cv_enet, s = "lambda.1se")
coef_custom <- coef(cv_enet, s = lambda_custom)

coef_min_df <- tibble(
  chemical = rownames(coef_min),
  coef_lambda_min = as.numeric(coef_min)
) %>%
  filter(chemical != "(Intercept)") %>%
  mutate(
    selected_lambda_min = coef_lambda_min != 0,
    abs_coef_lambda_min = abs(coef_lambda_min)
  )

coef_1se_df <- tibble(
  chemical = rownames(coef_1se),
  coef_lambda_1se = as.numeric(coef_1se)
) %>%
  filter(chemical != "(Intercept)") %>%
  mutate(
    selected_lambda_1se = coef_lambda_1se != 0,
    abs_coef_lambda_1se = abs(coef_lambda_1se)
  )

coef_custom_df <- tibble(
  chemical = rownames(coef_custom),
  coef_custom = as.numeric(coef_custom)
) %>%
  filter(chemical != "(Intercept)") %>%
  mutate(
    selected_custom = coef_custom != 0,
    abs_coef_custom = abs(coef_custom)
  )

enet_compare <- coef_min_df %>%
  select(chemical, coef_lambda_min, selected_lambda_min, 
        abs_coef_lambda_min) %>%
  left_join(
    coef_1se_df %>%
      select(chemical, coef_lambda_1se, 
            selected_lambda_1se, abs_coef_lambda_1se),
    by = "chemical"
  ) %>%
  left_join(
    coef_custom_df %>%
      select(chemical, coef_custom, selected_custom, abs_coef_custom),
    by = "chemical"
  ) %>%
  arrange(desc(abs_coef_lambda_min))

#enet_compare

```

```{r}
shortlist_min <- enet_compare %>%
  filter(selected_lambda_min) %>%
  arrange(desc(abs_coef_lambda_min))

shortlist_1se <- enet_compare %>%
  filter(selected_lambda_1se) %>%
  arrange(desc(abs_coef_lambda_1se))

shortlist_custom <- enet_compare %>%
  filter(selected_custom) %>%
  arrange(desc(abs_coef_custom))

```

```{r}
write.csv(enet_compare, "Elastic_Net_Chemical_Screening_Results.csv", 
        row.names = F)

```



Classical SIS screening (using pairwise Pearson correlation of X v/s y to screen):
```{r}
library(dccpp)
library(tidyverse)
sis_linear_tbl <- tibble(
  chemical = chem_cols,
  sis_corr = map_dbl(chem_cols, ~ cor(X[, .x], y, use = "complete.obs")),
  sis_abs_corr = abs(sis_corr)
) %>%
  arrange(desc(sis_abs_corr)) %>%
  mutate(
    sis_rank = row_number(),
    sis_top15 = sis_rank <= 15,
    sis_top20 = sis_rank <= 20,
    sis_top25 = sis_rank <= 25
  )

sis_dcorr_tbl <- tibble(
  chemical = chem_cols,
  dc_sis = map_dbl(chem_cols, ~ dccpp::dcor(X[, .x], y))
) %>%
  arrange(desc(dc_sis)) %>%
  mutate(
    dc_sis_rank = row_number(),
    dc_sis_top15 = dc_sis_rank <= 15,
    dc_sis_top20 = dc_sis_rank <= 20,
    dc_sis_top25 = dc_sis_rank <= 25
  )
```

Linear SIS uses pearson correlation so we are considering linear relationships, the other SIS considers
distance correlation and is basically looking at association, so it captures non-linear relationships.

Compiled Table with SIS results

```{r}
sis_compare <- sis_linear_tbl %>%
  select(
    chemical,
    sis_corr,
    sis_abs_corr,
    sis_rank,
    sis_top15,
    sis_top20,
    sis_top25
  ) %>%
  left_join(
    sis_dcorr_tbl %>%
      select(
        chemical,
        dc_sis,
        dc_sis_rank,
        dc_sis_top15,
        dc_sis_top20,
        dc_sis_top25
      ),
    by = "chemical"
  ) %>%
  mutate(
    mean_sis_rank = (sis_rank + dc_sis_rank) / 2,
    selected_by_both_top15 = sis_top15 & dc_sis_top15,
    selected_by_both_top20 = sis_top20 & dc_sis_top20,
    selected_by_both_top25 = sis_top25 & dc_sis_top25,
    selected_by_either_top15 = sis_top15 | dc_sis_top15,
    selected_by_either_top20 = sis_top20 | dc_sis_top20,
    selected_by_either_top25 = sis_top25 | dc_sis_top25
  ) %>%
  arrange(mean_sis_rank)

#head(sis_compare)

write.csv(sis_compare, "SIS_Chemical_Screening_Results.csv", row.names = FALSE)

```

PCA finds directions explaining variation in X.
PLS finds directions in X that are most related to Y.
Sparse PLS forces many variable weights to 0 so only selected predictors define the components.
Other than that, it works similar to PLS.

SPLS:
Tuning hyperparameters for SPLS:

Run this chunk to install mixOmics, you might need to do something different for a different version
of R:
```{r}
#install.packages("BiocManager")
#options(repos = BiocManager::repositories())
#getOption("repos") #- make sure this shows you all repos not just CRAN
#install.packages("ggrepel", repos = "https://cloud.r-project.org", type = "binary")
#BiocManager::install("mixOmics", ask = FALSE, update = FALSE)
#library(mixOmics)
```

```{r}
# options(repos = c(CRAN = "https://cloud.r-project.org"))
# install.packages("ggrepel")
# R.version.string
# sessionInfo()
```

```{r}
#install.packages("BiocManager")
#BiocManager::install("mixOmics")
library(mixOmics)
set.seed(42)
tune_spls_res <- tune.spls(
  X = as.matrix(X),
  Y = y,
  ncomp = 2,
  test.keepX = c(5, 10, 15, 20, 25),
  validation = "Mfold",
  folds = 5,
  nrepeat = 20,
  measure = "MSE",
  mode = "regression",
  progressBar = TRUE
)
```

Optimal hyperparameters:
```{r}
best_keepX <- tune_spls_res$choice.keepX
best_ncomp <- tune_spls_res$choice.ncomp

paste0("Optimal keepX: ", best_keepX, ", optimal ncomp: ", best_ncomp$ncomp)
```

KeepX = (15, 20), 20, ncomp = 2

2 PCs will be retained and 15 variables load on one PC and 20 on the other.
Variables are chosen through latent component optimization criterion, which is based on constructing
components that maximize covariance with the response while imposing sparsity.
There could be cross loadings.

I'd be inclined to think about the PCs as latent factors and less like just some projections.

Fitting the tuned models:
```{r}
best_keepX <- c(15, 20)
best_ncomp <- 2
X_mat <- as.matrix(X)
spls_fit <- spls(
  X = X_mat,
  Y = y,
  ncomp = best_ncomp,
  keepX = best_keepX,
  mode = "regression"
)


#Extracting selected variables
loadings_mat <- spls_fit$loadings$X

selected_list <- lapply(seq_len(best_ncomp), function(j) {
  rownames(mixOmics::selectVar(spls_fit, comp = j)$X$value)
})

names(selected_list) <- paste0("comp", seq_len(best_ncomp))

all_selected <- unique(unlist(selected_list))

spls_tbl <- tibble(
  chemical = rownames(loadings_mat),
  spls_selected = chemical %in% all_selected
)
```

Creating the results table:
```{r}
for (j in seq_len(ncol(loadings_mat))) {
  spls_tbl[[paste0("loading_comp", j)]] <- as.numeric(loadings_mat[, j])
}

loading_cols <- grep("^loading_comp", names(spls_tbl), value = TRUE)

spls_tbl <- spls_tbl %>%
  mutate(
    spls_importance = rowSums(across(all_of(loading_cols), ~ abs(.x)))
  ) %>%
  arrange(desc(spls_importance)) %>%
  mutate(
    spls_rank = row_number(),
    spls_top15 = spls_rank <= 15,
    spls_top20 = spls_rank <= 20,
    spls_top25 = spls_rank <= 25
  )

for (j in seq_len(best_ncomp)) {
  spls_tbl[[paste0("selected_comp", j)]] <- spls_tbl$chemical %in% selected_list[[j]]
}
```

Saving the results:
```{r}
write.csv(spls_tbl, "Sparse_PLS_Chemical_Screening_Results.csv", row.names = F)

```


Finally, we'll screen variables and give ranks based on XGBoost (ML model)

This is probably too complex of a solution but I hope when we are considering an ensemble of the
different screening techniques I'll give low weights to XGBoost rankings so that this does not sway
our final ranking too much.

```{r}
library(xgboost)
set.seed(42)
dtrain <- xgb.DMatrix(data = X_mat, label = y)

xgb_fit <- xgb.train(
  params = list(
    objective = "reg:squarederror",
    eval_metric = "rmse",
    eta = 0.05,
    max_depth = 4,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 0.8,
    lambda = 1,
    alpha = 0
  ),
  data = dtrain,
  nrounds = 200,
  verbose = 0
)

xgb_imp <- xgb.importance(
  feature_names = colnames(X),
  model = xgb_fit
)
```

xgb_gain is the entropy gain or how important the variable is.
xgb_cover shows how broadly the variable's splits affect observations.
xgb_frequency is how often the variable was used in splitting

Use the Gain to decide importance.

Creating a results table for XGBoost:
```{r}
xgb_tbl <- tibble(chemical = colnames(X)) %>%
  left_join(
    as_tibble(xgb_imp) %>%
      transmute(
        chemical = Feature,
        xgb_gain = Gain,
        xgb_cover = Cover,
        xgb_frequency = Frequency
      ),
    by = "chemical"
  ) %>%
  mutate(
    xgb_gain = replace_na(xgb_gain, 0),
    xgb_cover = replace_na(xgb_cover, 0),
    xgb_frequency = replace_na(xgb_frequency, 0)
  ) %>%
  arrange(desc(xgb_gain)) %>%
  mutate(
    xgb_rank = row_number(),
    xgb_top15 = xgb_rank <= 15,
    xgb_top20 = xgb_rank <= 20,
    xgb_top25 = xgb_rank <= 25
  )
write.csv(xgb_tbl, "XGBoost_Screening_Results.csv", row.names = FALSE)

```

Giving ranks to Elastic Net Screening:
```{r}
enet_compare <- enet_compare %>%
  mutate(
    enet_rank_lambda_min = min_rank(desc(abs_coef_lambda_min)),
    enet_rank_lambda_1se = min_rank(desc(abs_coef_lambda_1se)),
    enet_rank_custom = min_rank(desc(abs_coef_custom)),
    
    enet_top15_lambda_min = enet_rank_lambda_min <= 15,
    enet_top20_lambda_min = enet_rank_lambda_min <= 20,
    enet_top25_lambda_min = enet_rank_lambda_min <= 25,
    
    enet_top15_lambda_1se = enet_rank_lambda_1se <= 15,
    enet_top20_lambda_1se = enet_rank_lambda_1se <= 20,
    enet_top25_lambda_1se = enet_rank_lambda_1se <= 25,
    
    enet_top15_custom = enet_rank_custom <= 15,
    enet_top20_custom = enet_rank_custom <= 20,
    enet_top25_custom = enet_rank_custom <= 25
  )

enet_compare <- enet_compare %>%
  mutate(
    enet_rank_lambda_min = rank(-abs_coef_lambda_min, ties.method = "min"),
    enet_rank_lambda_1se = rank(-abs_coef_lambda_1se, ties.method = "min"),
    enet_rank_custom = rank(-abs_coef_custom, ties.method = "min")
  )

head(enet_compare)


```


Collating all Screening methods result to get one table:

```{r}
all_screening_tbl <- enet_compare %>%
  
  left_join(
    sis_compare,
    by = "chemical"
  ) %>%
  left_join(
    spls_tbl,
    by = "chemical"
  ) %>%
  left_join(
    xgb_tbl,
    by = "chemical"
  )

head(all_screening_tbl)
#Adding some summaries which will help me decide weights and validate assumptions
all_screening_tbl <- all_screening_tbl %>%
  mutate(
    mean_rank_all = rowMeans(
      cbind(
        sis_rank,
        dc_sis_rank,
        enet_rank_lambda_min,
        enet_rank_lambda_1se,
        enet_rank_custom,
        spls_rank,
        xgb_rank
      ),
      na.rm = TRUE
    ),
    
    n_top15_methods =
      sis_top15 + dc_sis_top15 +
      enet_top15_lambda_min + enet_top15_lambda_1se + enet_top15_custom +
      spls_top15 + xgb_top15,
    
    n_top20_methods =
      sis_top20 + dc_sis_top20 +
      enet_top20_lambda_min + enet_top20_lambda_1se + enet_top20_custom +
      spls_top20 + xgb_top20,
    
    n_top25_methods =
      sis_top25 + dc_sis_top25 +
      enet_top25_lambda_min + enet_top25_lambda_1se + enet_top25_custom +
      spls_top25 + xgb_top25
  ) %>%
  arrange(desc(n_top15_methods), mean_rank_all)


write.csv(all_screening_tbl, "Collated_Screening_Results.csv", row.names = FALSE)

```

Data Visualisation for comparing different methods:

```{r}
library(tidyverse)
library(ggplot2)
library(forcats)

heatmap_tbl <- all_screening_tbl %>%
  slice_min(mean_rank_all, n = 20) %>%
  dplyr::select(
    chemical,
    sis_rank,
    dc_sis_rank,
    enet_rank_lambda_min,
    enet_rank_lambda_1se,
    enet_rank_custom,
    spls_rank,
    xgb_rank
  ) %>%
  pivot_longer(
    cols = -chemical,
    names_to = "method",
    values_to = "rank"
  ) %>%
  mutate(
    method = factor(
      method,
      levels = c(
        "sis_rank",
        "dc_sis_rank",
        "enet_rank_lambda_min",
        "enet_rank_lambda_1se",
        "enet_rank_custom",
        "spls_rank",
        "xgb_rank"
      ),
      labels = c(
        "SIS",
        "DC-SIS",
        "ENet min",
        "ENet 1se",
        "ENet custom",
        "sPLS",
        "XGBoost"
      )
    )
  )

ggplot(heatmap_tbl,
       aes(x = method,
           y = fct_reorder(chemical, rank, .fun = min),
           fill = rank)) +
  geom_tile(color = "white") +
  geom_text(aes(label = rank), size = 3) +
  scale_fill_viridis_c(direction = -1) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Rank",
    title = "Top 20 chemicals across screening methods"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


Run each method multiple times, s.d., how stable is each method
rank 1: down weight by how unsure we are : weight= exp(-k*sd)
run this method iteratively (20-50x) and look at results and find s.d.


Bump Chart:
```{r}
library(tidyverse)
library(ggplot2)

bump_tbl <- all_screening_tbl %>%
  slice_min(mean_rank_all, n = 10) %>%
  dplyr::select(
    chemical,
    sis_rank,
    dc_sis_rank,
    enet_rank_lambda_min,
    enet_rank_lambda_1se,
    enet_rank_custom,
    spls_rank,
    xgb_rank
  ) %>%
  pivot_longer(
    cols = -chemical,
    names_to = "method",
    values_to = "rank"
  ) %>%
  mutate(
    method = factor(
      method,
      levels = c(
        "sis_rank",
        "dc_sis_rank",
        "enet_rank_lambda_min",
        "enet_rank_lambda_1se",
        "enet_rank_custom",
        "spls_rank",
        "xgb_rank"
      ),
      labels = c(
        "SIS",
        "DC-SIS",
        "ENet min",
        "ENet 1se",
        "ENet custom",
        "sPLS",
        "XGBoost"
      )
    )
  )

ggplot(bump_tbl,
       aes(x = method, y = rank, group = chemical, color = chemical)) +
  geom_line(size = 1.1, smooth = 8) +
  geom_point(size = 2) +
  scale_y_reverse(breaks = 1:20) +
  labs(
    x = NULL,
    y = "Rank",
    title = "Rank movement across screening methods"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


```



## Weighting Based on Stability of each method

```{r}

set.seed(2026)

n_runs <- 30
k_penalty <- 0.08
top_n_final <- 20

alpha_fixed <- 0.5
lambda_fixed <- exp(-2.8)

spls_ncomp_fixed <- 2
spls_keepX_fixed <- c(15, 20)

method_names <- c(
  "sis_rank",
  "dc_sis_rank",
  "enet_rank",
  "spls_rank",
  "xgb_rank"
)

```

We will used bootstrapped samples to see which are the most stable methods based on standard deviations and lower the standard deviation higher the weight in voting.

```{r}

rank_from_score <- function(scores, decreasing = TRUE) {
  if (decreasing) rank(-scores, ties.method = "average")
  else rank(scores, ties.method = "average")
}

#One Bootstrapped run
run_screening_once <- function(run_id, X, y, chem_cols) {
  
  set.seed(5000 + run_id)
  boot_idx <- sample(seq_len(nrow(X)), size = nrow(X), replace = TRUE)
  
  Xb <- X[boot_idx, , drop = FALSE]
  yb <- y[boot_idx]
  
  # Elastic Net (fixed alpha, fixed lambda)
  
  enet_fit_run <- glmnet(
    x = Xb,
    y = yb,
    alpha = 0.5,
    lambda = exp(-2.8),
    family = "gaussian",
    standardize = TRUE
  )
  
  coef_enet_run <- coef(enet_fit_run, s = lambda_fixed)
  
  enet_tbl_run <- tibble(
    chemical = rownames(coef_enet_run),
    enet_score = abs(as.numeric(coef_enet_run))
  ) %>%
    filter(chemical != "(Intercept)") %>%
    mutate(enet_rank = rank_from_score(enet_score, decreasing = TRUE)) %>%
    dplyr::select(chemical, enet_rank)
  
  # SIS Pearson
  
  sis_tbl_run <- tibble(
    chemical = chem_cols,
    sis_score = map_dbl(chem_cols, ~ abs(cor(Xb[, .x], yb, use = "complete.obs")))
  ) %>%
    mutate(sis_rank = rank_from_score(sis_score, decreasing = TRUE)) %>%
    dplyr::select(chemical, sis_rank)
  
  # DC-SIS

  dc_sis_tbl_run <- tibble(
    chemical = chem_cols,
    dc_sis_score = map_dbl(chem_cols, ~ dccpp::dcor(Xb[, .x], yb))
  ) %>%
    mutate(dc_sis_rank = rank_from_score(dc_sis_score, decreasing = TRUE)) %>%
    dplyr::select(chemical, dc_sis_rank)
  
  # sparse PLS 
  
  spls_fit_run <- spls(
    X = as.matrix(Xb),
    Y = yb,
    ncomp = 2,
    keepX = c(15, 20),
    mode = "regression"
  )
  
  loadings_run <- spls_fit_run$loadings$X
  
  spls_tbl_run <- tibble(
    chemical = rownames(loadings_run),
    spls_score = rowSums(abs(loadings_run))
  ) %>%
    mutate(spls_rank = rank_from_score(spls_score, decreasing = TRUE)) %>%
    dplyr::select(chemical, spls_rank)
 
  # XGBoost (fixed parameters)
  
  dtrain_run <- xgb.DMatrix(data = as.matrix(Xb), label = yb)
  
  xgb_fit_run <- xgb.train(
    params = list(
      objective = "reg:squarederror",
      eval_metric = "rmse",
      eta = 0.05,
      max_depth = 4,
      min_child_weight = 3,
      subsample = 0.8,
      colsample_bytree = 0.8,
      lambda = 1,
      alpha = 0
    ),
    data = dtrain_run,
    nrounds = 200,
    verbose = 0
  )
  
  xgb_imp_run <- xgb.importance(
    feature_names = colnames(Xb),
    model = xgb_fit_run
  )
  
  xgb_tbl_run <- tibble(chemical = colnames(Xb)) %>%
    left_join(
      as_tibble(xgb_imp_run) %>%
        transmute(chemical = Feature, xgb_score = Gain),
      by = "chemical"
    ) %>%
    mutate(xgb_score = replace_na(xgb_score, 0)) %>%
    mutate(xgb_rank = rank_from_score(xgb_score, decreasing = TRUE)) %>%
    dplyr::select(chemical, xgb_rank)
  
  
  # Join all ranks
  
  sis_tbl_run %>%
    left_join(dc_sis_tbl_run, by = "chemical") %>%
    left_join(enet_tbl_run, by = "chemical") %>%
    left_join(spls_tbl_run, by = "chemical") %>%
    left_join(xgb_tbl_run, by = "chemical") %>%
    mutate(run = run_id)
}
```

Executing bootstrapped runs

```{r}

# REPEATED BOOTSTRAP RUNS

rank_runs <- map_dfr(1:n_runs, ~ run_screening_once(.x, X, y, chem_cols))

write.csv(rank_runs, "Repeated_Bootstrap_Method_Ranks.csv", row.names = FALSE)


# METHOD STABILITY
# Mean SD of chemical ranks within each method

method_stability <- map_dfr(method_names, function(m) {
  rank_runs %>%
    group_by(chemical) %>%
    summarise(rank_sd = sd(.data[[m]], na.rm = TRUE), .groups = "drop") %>%
    summarise(
      method = m,
      mean_rank_sd = mean(rank_sd, na.rm = TRUE),
      median_rank_sd = median(rank_sd, na.rm = TRUE)
    )
})

method_stability <- method_stability %>%
  mutate(
    raw_weight = exp(-k_penalty * mean_rank_sd),
    weight = raw_weight / sum(raw_weight)
  ) %>%
  arrange(desc(weight))

write.csv(method_stability, "Method_Stability_Weights.csv", row.names = FALSE)


# AVERAGE RANKS ACROSS BOOTSTRAP RUNS

avg_rank_tbl <- rank_runs %>%
  group_by(chemical) %>%
  summarise(
    sis_rank = mean(sis_rank, na.rm = TRUE),
    dc_sis_rank = mean(dc_sis_rank, na.rm = TRUE),
    enet_rank = mean(enet_rank, na.rm = TRUE),
    spls_rank = mean(spls_rank, na.rm = TRUE),
    xgb_rank = mean(xgb_rank, na.rm = TRUE),
    .groups = "drop"
  )


# WEIGHTED COMPOSITE RANK

weight_vec <- method_stability$weight
names(weight_vec) <- method_stability$method

final_rank_tbl <- avg_rank_tbl %>%
  mutate(
    composite_score =
      weight_vec["sis_rank"] * sis_rank +
      weight_vec["dc_sis_rank"] * dc_sis_rank +
      weight_vec["enet_rank"] * enet_rank +
      weight_vec["spls_rank"] * spls_rank +
      weight_vec["xgb_rank"] * xgb_rank
  ) %>%
  arrange(composite_score) %>%
  mutate(
    composite_rank = row_number()
  )

top20_final_tbl <- final_rank_tbl %>%
  slice_head(n = top_n_final)


# FINAL TABLE WITH INDIVIDUAL RANKS + COMPOSITE

final_table <- final_rank_tbl %>%
  dplyr::select(
    chemical,
    sis_rank,
    dc_sis_rank,
    enet_rank,
    spls_rank,
    xgb_rank,
    composite_score,
    composite_rank
  )

write.csv(final_table, "Final_Composite_Ranking_Table.csv", row.names = FALSE)
write.csv(top20_final_tbl, "Top20_Composite_Chemicals.csv", row.names = FALSE)


# METHOD WEIGHTS TABLE

method_weights_table <- method_stability %>%
  dplyr::select(method, mean_rank_sd, median_rank_sd, weight)

write.csv(method_weights_table, "Method_Weights_Table.csv", row.names = FALSE)
```

```{r}

top20_final_tbl
method_weights_table
```


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(scales)

# top20_final_tbl should contain:
# chemical, sis_rank, dc_sis_rank, enet_rank, spls_rank, xgb_rank, composite_score
#
# method_weights_table should contain:
# method, mean_rank_sd, median_rank_sd, weight

# Order chemicals by composite score: lower = better
plot_dat <- top20_final_tbl %>%
  arrange(composite_score) %>%
  mutate(chemical = factor(chemical, levels = chemical))

# Long format for method-specific rank plot
rank_long <- plot_dat %>%
  dplyr::select(chemical, sis_rank, dc_sis_rank, enet_rank, spls_rank, xgb_rank) %>%
  pivot_longer(
    cols = -chemical,
    names_to = "method",
    values_to = "rank"
  ) %>%
  mutate(
    method = recode(
      method,
      sis_rank = "SIS",
      dc_sis_rank = "DC-SIS",
      enet_rank = "Elastic Net",
      spls_rank = "sPLS",
      xgb_rank = "XGBoost"
    )
  )

# 1) Multi-line plot of ranks across methods
p1 <- ggplot(rank_long, aes(x = chemical, y = rank, color = method, group = method)) +
  geom_line(linewidth = 0.9, alpha = 0.9) +
  geom_point(size = 2) +
  scale_y_reverse(breaks = pretty_breaks(n = 8)) +
  labs(
    title = "Method-Specific Rankings of Top Chemicals",
    x = "Chemical",
    y = "Rank (1 = best)",
    color = "Method"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )

# Save first plot
ggsave(
  filename = "top20_final_tbl.png",
  plot = p1,
  width = 11,
  height = 6.5,
  dpi = 300
)

# 2) Single-line plot of composite score
p2 <- ggplot(plot_dat, aes(x = chemical, y = composite_score, group = 1)) +
  geom_line(color = "#2C3E50", linewidth = 1) +
  geom_point(color = "#C0392B", size = 2.5) +
  scale_y_reverse() +
  labs(
    title = "Composite Ranking of Top Chemicals",
    x = "Chemical",
    y = "Composite Rank"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

# Save second plot
ggsave(
  filename = "method_weights_table.png",
  plot = p2,
  width = 11,
  height = 6,
  dpi = 300
)
```
