---
title: "Word_Cloud_Each_Sample"
output: pdf_document
date: "2026-02-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```




Data Under Consideration:
```{r, echo=FALSE}
library(dplyr)
pb <- read.csv("Data/PB.csv") #Change file location based on where your file is

pb_feedback <- pb[, c("Unique_Panelist_ID", "Sample", "Comments_aroma_liking", "Comments_aroma_disliking", "Aroma_Liking_A")] 

pb_concat_feedback <- pb_feedback %>% 
  distinct()

head(pb_concat_feedback, 10)
```



Cleaning the text data:

1. Rename columns for easier coding  

2. Recode things that look like NAs or null values to NA  

3. Drop the rows where both positive and negative reviews are NA or null  

4. Unpivot the data each customer x sample combination will have 2 rows one for positive and one for negative review (need this for the way I created the wordclouds)

5. Tokenize the words and remove stop words

6. Count the frequency of words (by sample and sentiment) and then calculate TF-Idf (Term frequency Inverse Document frequency)


TF-Idf penalizes generic words that are frequently occuring in reviews for other samples


TF = Probability (proportion) of seeing a word in a document (one sample X sentiment combination)
IDF = log(N/df_w), 

N = the total number of documents (total sample x sentiment combinations),

df_w = number of documents containing the given word

TF-idf = TF * IDF = P(observing given word in your document) * how often the word appears overall

Intuituion: We are penalizing common words, we want words that correlate with specific cheese to have high values.



```{r, include=FALSE}
library(tidyverse)
library(tidytext)
library(wordcloud)
library(RColorBrewer)
library(tm)


#Renaming columns:
df <- pb_concat_feedback
df <- df %>%
  rename(
    positive_text = Comments_aroma_liking,
    negative_text = Comments_aroma_disliking
  )

missing_like <- function(x) {
  x <- trimws(tolower(x))
  x[x %in% c("n/a", "na", "nan", "null", "nil", "none", "missing", 
               "not available", "")] <- NA
  x
}

#Replace NA like Values with NA
df <- df %>%
  mutate(across(c(positive_text, negative_text), missing_like))

#Rows where both positive and negative are NA
rows_both_na <- df %>%
  filter(is.na(positive_text) & is.na(negative_text))

#Drop the rows where both are NA
df <- df %>%
  filter(!(is.na(positive_text) & is.na(negative_text)))

#If Positive review is NA but there's a negative review, impute with "terrible" because I am assuming this means the customer didn't find anything positive about the cheese
#If Negative review is NA but there's a positive review, impute with "amazing" because I am assuming this means the customer didn't find anything negative about the cheese
df <- df %>%
  mutate(
    positive_text = ifelse(
      is.na(positive_text) & !is.na(negative_text),
      "terrible",
      positive_text
    ),
    negative_text = ifelse(
      is.na(negative_text) & !is.na(positive_text),
      "amazing",
      negative_text
    )
  )
```



Unpivot for comparisons and easier computations, remove stop words, tokenize and calculate frequency
```{r, include=FALSE}
#Putting data in correct format - unpivoting
df_long <- df %>%
  pivot_longer(
    cols = c(positive_text, negative_text),
    names_to = "sentiment",
    values_to = "text"
  ) %>%
  mutate(sentiment = ifelse(sentiment == "positive_text", "positive", "negative"))
#Text cleaning and tokenization
data("stop_words")

tokens <- df_long %>%
  mutate(text = str_to_lower(text)) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word") %>%#Removing stop words like is, are etc.
  filter(!str_detect(word, "^[0-9]+$"))


#Word counts by sample and sentiment - counting tokenized words
word_counts <- tokens %>%
  count(Sample, sentiment, word, sort = TRUE)

#Tf-idf for each sample - use this for wordclouds, penalizes generic words
#doc_id = paste(Sample, sentiment)
tfidf <- word_counts %>%
  unite(doc_id, Sample, sentiment, remove = FALSE) %>%
  bind_tf_idf(word, doc_id, n)
```



Function to create wordclouds given Sample and whether you want positive or negative review takes ch and pos as input. 

ch: Sample Name (A, B, C,...)

pos: positive or negative. 


Set ch = sample (A, B, C,....) and pos = "positive" if you'd like a wordcloud for positive review and "negative" if you'd like one for negative reviews.

```{r, include=FALSE}
word_cloud_review <- function(ch, pos){
  op <- par(mar = c(5, 4, 4, 2))  # increase bottom margin
  
  tfidf %>%
  filter(Sample == ch, sentiment == pos) %>%
  with({wordcloud(
    words = word,
    freq = tf_idf,
    max.words = 100,#Reduce this if you think too many unnecassary words are showing up
    colors = brewer.pal(8, "Dark2"),
    
  )
    mtext(
        text = paste(str_to_upper(pos), "REVIEWS -", ch),
        side = 1,      # bottom
        line = 4,      # distance from plot
        cex = 1.1
      )}
  )
  par(op)
}
```



Sample A:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("A", "positive")
```



Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("A", "negative")
```



Sample B:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("B", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("B", "negative")
```


Sample C:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("C", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("C", "negative")
```



Sample D:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("D", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("D", "negative")
```



Sample E:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("E", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("E", "negative")
```



Sample F:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("F", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("F", "negative")
```



Sample G:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("G", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("G", "negative")
```



Sample H:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("H", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("H", "negative")
```



Sample I:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("I", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("I", "negative")
```



Sample J:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("J", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("J", "negative")
```




Sample K:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("K", "positive")
```
Sample K is the most liked cheese by Median Aroma Liking

Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("K", "negative")
```



Sample L:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("L", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("L", "negative")
```



Sample M:

Positive Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("M", "positive")
```


Negative Reviews:
```{r, echo=FALSE}
#Positive review
word_cloud_review("M", "negative")
```